---
title: "HW 02"
author: "Weston Scott"
format: 
  html:
    embed-resources: true
toc: true
editor:
  render-on-save: true
execute:
  warning: false
# bibliography: references.bib
---

```{r library_setup, include=FALSE}
if (!require("pacman")) 
    install.packages("pacman")

# Use pacman::p_load to install and load CRAN packages
pacman::p_load(
    dplyr,
    forcats,
    fs,
    ggridges,
    janitor,
    lubridate,
    paste,
    readr,
    scales,
    stringr,
    tidyverse,
    viridis
)

# Handle GitHub package separately
if (!require("dsbox")) {
      # Install devtools if not present
      if (!require("devtools")) 
          install.packages("devtools")
devtools::install_github("tidyverse/dsbox")
      library(dsbox)
}
```

```{r set_uniform_displays, include=FALSE}
# set theme for ggplot2
# ggplot2::theme_set(ggplot2::theme_minimal(base_size = 10))

# # set width of code output
# options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7, # 7" width
  fig.asp = 0.618, # the golden ratio
  fig.retina = 3, # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300 # higher dpi, sharper image
)
```

## 1 - A new day, a new plot, a new geom

```{r load_data_problem_1}
edibnb <- dsbox::edibnb
glimpse(edibnb)
summary(edibnb)

edibnb <- edibnb |>
    mutate(
        neighbourhood = fct_reorder(
            neighbourhood, 
            review_scores_rating, 
            .fun = median)
    ) |>
    filter(!is.na(neighbourhood))
```

```{r ridgeline_plot_problem_1}
ggplot(
    data = edibnb, 
    aes(
        x = review_scores_rating, 
        y = neighbourhood, 
        fill = neighbourhood
    )
) +
    
geom_density_ridges(
    scale = 2,
    rel_min_height = 0.01,
    legend.show = FALSE,
    alpha = 0.8
) +

scale_fill_viridis_d(
    option = "C", 
    begin = 0.1, 
    end = 0.9
) +

scale_y_discrete(expand = c(0, 0)) + 
scale_x_continuous(
    expand = c(0, 0.25),
    limits = c(90, 100)
) + 

coord_cartesian(clip = "off") +
labs(
    title = "Problem 1 - Ridgeline plot", 
    subtitle = "Airbnb listings: Edinburgh, Scotland",
    x = "Review Score Ratings",
    y = "Edinburgh\nNeighborhoods", 
    caption = "Source: Opensource dataset dsbox::edibnb"
) +

theme_ridges() +
theme(legend.position = "none")
```

### Intepretation

The `ridgeline` plot above visualizes the distribution of Airbnb review scores across different Edinburgh neighborhoods, ordered by their respective median review scores. The neighborhoods with the highest overall reviews appear at the top of the plot with a descending order down the plot to the neighborhoods with the lowest review scores. Most review scores cluster tightly between 90 and 100 (x-axis), suggesting generally positive experiences overall in the set of reviews. However, some neighborhoods display broader distributions or lower medians. A broader distribution indicates that the reviews have a larger spread along the review spectrum.

## 2 - Foreign Connected PACs

```{r load_data_problem_2}
# get a list of files with "Foreign Connected PAC" in their names
list_of_files <- dir_ls(path = "data", regexp = "Foreign Connected PAC")

# read all files and row bind them
# keeping track of the file name in a new column called year
pac <- read_csv(list_of_files, id = "year")
glimpse(pac)
```

```{r extend_data_problem_2}
pac <- pac |>
    clean_names() |>
    separate(
        country_of_origin_parent_company,
        into = c("country", "parent_company"),
        sep = "/", 
        remove = TRUE) |>

    mutate(
        year = str_extract(year, "\\d{4}-\\d{4}"),
        year = str_extract(year, "\\d{4}$"),
        year = as.integer(year)
    ) |>

    select(-total)

pac

pac <- pac |>
    pivot_longer(
        cols = c(dems, repubs),
        names_to = "party",
        values_to = "amount"
    ) |>

    mutate(
        amount = str_remove(amount, "\\$"),
        amount = as.integer(amount)
    )

pac
```

```{r summarize_party_amounts_problem_2}
uk_spending <- pac |>
    filter(country == "UK") |>
    group_by(year, party) |>

    summarise(
        totals = sum(amount[amount != 0],
                     na.rm = TRUE), 
        .groups = "drop") |>

    arrange(year, party)

uk_spending
```

```{r plot_uk_spending_problem_2}
ggplot(data = uk_spending, 
       aes(x = year, y = totals / 1e6, color = party)) +

geom_line() +
scale_color_manual(values = c("repubs" = "red", 
                              "dems" = "blue"),
                  labels = c("Democrats", "Republicans")) +
scale_x_continuous(breaks = seq(1996, 2022, by = 4),
                   labels = seq(1996, 2022, by = 4)) +
scale_y_continuous(labels = function(x) paste0("$", round(x), "M")) +

labs(
    title = "Contributions to US political parties from UK-connected PACs",
    color = "Party",
    x = "Year",
    y = "Total amount",
    caption = "Source: OpenSecrets.org"
) +
theme_minimal() +
theme(
    axis.title.y = element_text(margin = margin(t = 0, r = 0, 
                                                b = 0, l = 0),
                                hjust = 0),
    axis.title.x = element_text(margin = margin(t = 0, r = 0, 
                                                b = 0, l = 0), 
                                hjust = 0),
    legend.position = c(0.87, 0.15)
)

```

```{r plot_other_country_spending_problem_2}
swiss_spending <- pac |>
    filter(country == "Switzerland") |>
    group_by(year, party) |>

    summarise(
        totals = sum(amount[amount != 0],
                     na.rm = TRUE), 
        .groups = "drop") |>

    arrange(year, party)


ggplot(data = swiss_spending, 
       aes(x = year, y = totals / 1e6, color = party)) +

geom_line() +
scale_color_manual(values = c("repubs" = "red", 
                              "dems" = "blue"),
                  labels = c("Democrats", "Republicans")) +
scale_x_continuous(breaks = seq(1996, 2022, by = 4),
                   labels = seq(1996, 2022, by = 4)) +
scale_y_continuous(labels = function(x) paste0("$", round(x), "M")) +

labs(
    title = "Contributions to US political parties from Swiss-connected PACs",
    color = "Party",
    x = "Year",
    y = "Total amount",
    caption = "Source: OpenSecrets.org"
) +
theme_minimal() +
theme(
    axis.title.y = element_text(margin = margin(t = 0, r = 0, 
                                                b = 0, l = 0), 
                                hjust = 0),
    axis.title.x = element_text(margin = margin(t = 0, r = 0, 
                                                b = 0, l = 0), 
                                hjust = 0),
    legend.position = c(0.87, 0.15)
)

```

### Intepretation

Contributions from Swiss-connected PACs to U.S. political parties have grown since the year 2000. The peak around key election years. The data shows a clear preference for Republican candidates, especially from 2008 onward. This might reflect Swiss alignment of ideologies or policies with Republican platforms. In contrast, Democratic contributions also grew, though they remained more modest and stable over the yearly span of this dataset.

## 3 - Median housing prices in the US

```{r load_data_problem_3}
median_housing <- read_csv("data/median-housing.csv")

median_housing <- median_housing |>
  rename(date = DATE) |>
  rename(price = MSPUS)
glimpse(median_housing)

recessions <- read_csv("data/recessions.csv")
glimpse(recessions)
```

```{r plot_median_prices_problem_3}
ggplot(data = median_housing, 
       aes(x = date, y = price)) +

geom_line(color = "darkblue") +
scale_x_date(breaks = seq(as.Date("1960-01-01"), 
                          as.Date("2020-01-01"), 
                          by = "5 years"),
             labels = date_format("%Y")) +

scale_y_continuous(breaks = seq(0, 400000, by = 40000),
                   labels = label_number(accuracy = 1, 
                                         big.mark = ",")) +

labs(
    title = "Median sales prices of houses sold in the United States",
    subtitle = "Not seasonally adjusted",
    x = NULL,
    y = "Dollars",
    caption = "Source: Census; HUD"
) +
theme_minimal() +
theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank()
)
```

```{r plot_median_prices_recessions_problem_3}
recessions <- recessions |>
    mutate(
        is_recess = if_else(Peak >= as.Date("1963-01-01", 
                                            format = "%Y-%m-%d") & 
                            Trough <= as.Date("2021-04-01", 
                                              format = "%Y-%m-%d"),
                            TRUE, FALSE)
    ) |>
    filter(is_recess == TRUE)
glimpse(recessions)

ggplot(data = median_housing, 
       aes(x = date, 
           y = price)) +

geom_rect(
    data = recessions,
    aes(
        xmin = as.Date(Peak), 
        xmax = as.Date(Trough),
        ymin = -Inf, 
        ymax = Inf,
        y = NULL,
        x = NULL
    ), 
    fill = "cornsilk3"
    ) +

geom_line(color = "darkblue") +
scale_x_date(breaks = seq(as.Date("1960-01-01"), 
                          as.Date("2020-01-01"), 
                          by = "5 years"),
             labels = date_format("%Y")) +

scale_y_continuous(breaks = seq(0, 400000, by = 40000),
                   labels = label_number(accuracy = 1, big.mark = ",")) +

labs(
    title = "Median sales prices of houses sold in the United States",
    subtitle = "Not seasonally adjusted",
    x = NULL,
    y = "Dollars",
    caption = "Shaded areas indicate U.S. recessions\nSource: Census; HUD"
) +
theme_minimal() +
theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank()
)
```

## 4 - Expect More. Plot More.

## 5 - Mirror, mirror on the wall, who’s the ugliest of them all?